Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging

### 2022-12-04 script/manuscript_MEE.Rmd
some writing, need to review how many seqs are classified, it is in excess of raw

---
title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
author: |
  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
abstract: |
  an abstract will be written to fill this space. it will have from 4-6 numbered sections.
output:
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: true
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

  Recently barcoding, and metabarcoding, have shown considerable promise in all Kingdoms of life. For example …. . With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), with most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna). Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant) Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.

  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the spatial extent of the study region. While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, (typically Government) survey work, and citizen science (e.g. iNaturalist), from a domain exceeding the study area. To these candidate species modelling approaches, such as logistic regression, may be used to identify distances under which taxa warrant further exploration. To these candidate species, we generate species distribution models (SDM's), which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results. This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers. Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.

@@

```{r eval = F}
toseq <- read.csv('../data/Benkendorf_combos_fixed.csv') |> 
  filter(!str_detect(extract_id, '[aA-zZ]')) %>% nrow()

@@

XX corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(toseq)` corbiculae samples were sequenced, XX reads were recovered from sequencing. After trimming XX sequences remained. `r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST. The number of reads per sequence varied widely (min = , $\bar{x}$, max = ), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.

Kraken2 was able to identify the species richness of pollen samples ($\bar{x}$ = , min = , max = ). Bracken was able to estimate the relative abundance of pollen grains in each sample (max % of any species, $\bar{x}$ % of all species, min % trace amounts detected). BLAST, the most thorough sequence alignment algorithm...

```{r, eval = F}
rm(toseq)
```

@@

>>>>

---
title: Plant Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging
  Patterns of Queen Bumble Bees in the Southern Rocky Mountains, U.S.A.
author: |
  Reed Clark Benkendorf^[Correspondence: reedbenkendorf2021@u.northwestern.edu] $^1$$^,$$^2$, Jane E. Ogilvie $^3$, Emily J. Woodworth $^1$$^,$$^2$, Sophie Taddeo $^1$$^,$$^2$, Paul J. CaraDonna $^1$$^,$$^2$$^,$$^3$, Jeremie B. Fant $^1$$^,$$^2$ | $^1$Chicago Botanic Garden, 1000 Lake Cook Road, Glencoe, Illinois 60022 USA,  $^2$Plant Biology and Conservation, Northwestern University, Evanston, IL 60208, USA, $^3$ Rocky Mountain Biological Laboratory, P.O. Box 519, Crested Butte, CO 81224, USA
abstract: |
  an abstract will be written to fill this space. it will have from 4-6 numbered sections. like 1) A very broad overall identifying organisms is hard and we dont have time. 2) to try and figure this out we used whoa molecular stuff and holy shit amounts of computing 3) Our results showed we were able to do some cool things 4) you should try this out, because it was pretty cool and helped us perform natural history on steroids.
output:
  pdf_document: default
  toc: no
  word_document: default
csl: "../citations/methods-in-ecology-and-evolution.csl"
bibliography: ../citations/citations.bib
link-citations: yes
fig_caption: yes
always_allow_html: true
header-includes:
- \usepackage{endfloat}
- \usepackage{setspace}\doublespacing
- \usepackage{lineno}
- \linenumbers
---

@@

  Recently barcoding, and metabarcoding, have shown considerable promise in all Kingdoms of life. For example …. . With plants the identification of members of certain clades has been quite successful, whereas with others results have been elusive (@liu2014identification, @china2011comparative), while most applications laying along this spectrum (@li2015plant, @kress2007two, @cbol2009dna). Particular challenges with the utilization of high-copy number sequences are associated with their rates of divergence, gene tree conflict, and hybridization (@coissac2016barcodes, @fazekas2009plant) Herein we have resolved major components of the problems of identifying plant material without diagnostic morphological character states using the Angiosperms353 (A353) Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering.

  Our foundation for increasing the quality of metabarcoding results in plants is reducing the number of possible plant species candidates by generating user selected sequence databases relevant to the spatial extent of the study region. While there are numerous possible approaches for this process, we achieve the selection of possible plant candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g. iNaturalist), from a domain exceeding the study area. To these candidate species, modelling approaches - such as logistic regression, may be used to identify distances under which taxa warrant further exploration. To these candidate species, we generate species distribution models (SDM's), which indicate the probability of suitable habitat in a domain, and base the inclusion of these taxa, or representative congeners, upon these results. This approach has the additional benefit of greatly reducing the size of a sequence database, which allows for the usage of genomic size data on personal computers. Moreover, as most next-generation sequence data is deposited as raw-sequence reads, from a processing perspective, it is essential to reduce the candidate species via an approach as such.

@@

```{sh Get number of raw reads for metagenome samples, eval = F}

cd ~/sequenceData/Original_zip
files=$(find -type f -name "[0-9][0-9]_S[0-9]*_paired.fastq.gz")

# seqs per each sample , need divide by 4 on import
wc -l  $files > '/hdd/yun master/MEE_manuscript/data/raw_reads_metagenome_seq.txt'
```

```{r eval = F}

raw_seqs <- read.delim('../data/raw_reads_metagenome_seq.txt', header = F) %>% 
  mutate(V1 = str_trim(V1)) %>% 
  separate(V1, c('seq', 'sampleFR'), sep = ' ') %>% 
  mutate(seq = round(as.numeric(seq)/4),
         sample = str_extract(sampleFR, '.*L002_')) %>% 
  filter(sampleFR != 'total') %>% 
  group_by(sample) %>% 
  mutate(totalReads = sum(seq)) %>% 
  distinct(totalReads) %>% 
  pull(totalReads)

@@

54 corbiculae loads had DNA extracted and underwent hyb-seq, a total of `r toString(length(raw_seqs))` corbiculae samples were sequenced, `r toString(sum(raw_seqs))` reads were recovered from sequencing. The number of reads per sequence varied widely (min = `r toString(min(raw_seqs))`, $\bar{x}$ = `r toString(mean(raw_seqs))`, Mdn = `r toString(median(raw_seqs))`, max = `r toString(max(raw_seqs))`), the number of loci recovered per sample ranged from (min = , max = ), and the number of reads from loci also ranged widely () **APPENDIX X**.. After trimming XX sequences remained. `r toString(sum(krak[krak$Classified == 'C',]$n))` reads were matched using Kraken, of the reads classified by Kraken `r toString(sum(brack$reads))` reads were matched using Bracken, of the reads classified by Kraken `r toString(sum(blast$Total_seqs))` reads were matched using BLAST.

Based upon subjective review of the three classifiers, BLAST was chosen as the classification method which yielded the most probable results, and it's values were used for all analyses.

```{r, eval = F}
rm(toseq, krak, brack, blast)
```

@@

Paul CaraDonna https://orcid.org/0000-0003-3517-9090
Jeremie Fant https://orcid.org/0000-0001-9276-1111
Jane Ogilvie https://orcid.org/0000-0001-8546-0417
Sophie Taddeo https://orcid.org/0000-0002-7789-1417
